### 2017-11-21 PosterNIPS2017.tex
Poster

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% baposter Landscape Poster
% LaTeX Template
% Version 1.0 (11/06/13)
%
% baposter Class Created by:
% Brian Amberg (baposter@brian-amberg.de)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[portrait,a0paper,fontscale=0.35]{baposter} % Adjust the font scale/size here : fontscale=0.285
%\usepackage{textpos}
\usepackage[frenchb]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{graphicx} % Required for including images
%\graphicspath{{Figures/}} % Directory in which figures are stored
\usepackage{amsmath} % For typesetting math
\usepackage{amsthm}
\usepackage{amssymb} % Adds new symbols to be used in math mode
\usepackage{booktabs} % Top and bottom rules for tables
\usepackage{enumitem} % Used to reduce itemize/enumerate spacing
\usepackage{palatino} % Use the Palatino font
\usepackage[font=small,labelfont=bf]{caption} % Required for specifying captions to tables and figures
\usepackage{multicol} % Required for multiple columns
\usepackage{tikz} % Required for flow chart
\usetikzlibrary{shapes,arrows} % Tikz libraries required for the flow chart in the template
%\usepackage{dsfont}
%\usepackage{listings}
\usepackage{subfigure}
\usepackage{color}
%\usepackage{nth}
%\usepackage{a4wide}
\usepackage{graphicx,color}
\usepackage{array}
%\usepackage[nottoc, notlof, notlot]{tocbibind}
\usepackage{subfigure}
\usetikzlibrary{arrows,shapes,positioning,calc}
%\definecolor{CPU}{RGB}{30, 167, 223}
%\definecolor{CPU}{RGB}{8, 138, 104}
\definecolor{gris}{gray}{0.9}
\usepackage{geometry}
\geometry{left=0.5cm,right=0.5cm,bottom=0.5cm,top=0.5cm,centering}
%\usepackage{subfig}
\usepackage{float}
\usepackage{multicol}
\usepackage{enumitem,rotating}
\usepackage{bbm}
\usepackage{tcolorbox}

\newcommand{\sidecap}[1]{ {\begin{sideways}\parbox{0.2\textwidth}{\centering #1}\end{sideways}} }

\theoremstyle{plain}
\newtheorem*{thm}{Theorem}
%\theoremstyle{definition}
%\newtheorem*{defi}[thm]{Definition}
\theoremstyle{plain}
\newtheorem*{prop}{Properties}
\theoremstyle{plain}
\newtheorem*{hyp}{Assumption}
\theoremstyle{plain}
\newtheorem*{proposition}{Proposition}

\setlength{\columnsep}{1.5em} % Slightly increase the space between columns
\setlength{\columnseprule}{0mm} % No horizontal rule between columns

\definecolor{lightred}{rgb}{0.5 0 0}

\newcommand{\compresslist}{ % Define a command to reduce spacing within itemize/enumerate environments, this is used right after \begin{itemize} or \begin{enumerate}
\setlength{\itemsep}{1pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}
}

\definecolor{CPU}{RGB}{0, 153, 125} % Defines the color used for content box headers

% Math operators
\newcommand{\scal}[2]{\left\langle #1 , #2 \right\rangle}
\DeclareMathOperator{\IR}{\mathbb{R}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\One}{\mathbbm{1}}
\DeclareMathOperator{\Ccal}{\mathcal{C}}
\DeclareMathOperator{\logsumexp}{logsumexp}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\KL}{KL}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\renewcommand{\epsilon}{\varepsilon}

\begin{document}

\begin{poster}
{
background=plain,
headershade=plain,
headerborder=closed, % Adds a border around the header of content boxes
colspacing=0.8em, % Column spacing
bgColorOne=gris, % Background color for the gradient on the left side of the poster
bgColorTwo=gris, % Background color for the gradient on the right side of the poster
borderColor=CPU, % Border color
headerColorOne=CPU, % Background color for the header in the content boxes (left side)
headerColorTwo=CPU, % Background color for the header in the content boxes (right side)
headerFontColor=white, % Text color for the header text in the content boxes
boxColorOne=white, % Background color of the content boxes
textborder=roundedsmall, % Format of the border around content boxes, can be: none, bars, coils, triangles, rectangle, rounded, roundedsmall, roundedright or faded
eyecatcher=true, % Set to false for ignoring the left logo in the title and move the title left
headerheight=0.10\textheight, % Height of the header
headershape=rounded, % Specify the rounded corner in the content box headers, can be: rectangle, small-rounded, roundedright, roundedleft or rounded
headerfont=\Large\bf\textsc, % Large, bold and sans serif font in the headers of content boxes
%textfont={\setlength{\parindent}{1.5em}}, % Uncomment for paragraph indentation
linewidth=1.8pt, % Width of the border lines around content boxes
columns=2,
}
%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------
%
{\includegraphics[height=5em]{logo_ens_psl.png}} % First university/lab logo on the left
{\bf Overrelaxed Sinkhorn--Knopp algorithm\\ for regularized optimal transport\vspace{0.2em}} 
{{Alexis Thibault, L\'enaïc Chizat, Charles Dossal and  Nicolas Papadakis}} % Author names and institution
{\includegraphics[height=5em]{logo_imb.png}} % Second university/lab logo on the right

%----------------------------------------------------------------------------------------
%	RESEARCH
%----------------------------------------------------------------------------------------

% Introduction

\headerbox{Introduction}{name=intro,column=0}{
	Optimal Transport is an efficient and flexible tool to compare two probability distributions. It has been popularized in the computer vision community in the context of discrete histograms [Rubner, 2000]. The introduction of entropic regularization in [Cuturi, 2013] has made possible the use of the fast Sinkhorn--Knopp algorithm (SK) [Sinkhorn, 1964c] scaling with high dimensional data.\\
	\begin{minipage}{\textwidth}
		\centering
		\begin{tcolorbox}[width=0.99\textwidth,/tcb/size=small, /tcb/top=2.0mm]
			\centering
			\begin{tabular}{c c c c}
				\includegraphics[height=3cm]{1d_interp_mu} &
				\includegraphics[height=3cm]{1d_interp_0_2} &
				\includegraphics[height=3cm]{1d_interp_2} &
				\includegraphics[height=3cm]{1d_interp_18} \\
				&
				$\epsilon = 5$ &
				$\epsilon = 0.5$ &
				$\epsilon = 0.02$
			\end{tabular}
			\captionof{figure}{Regularized optimal transport maps for decreasing regularization parameter $\epsilon$.} \label{fig:1d_ot}
		\end{tcolorbox}
	\end{minipage}
	For small regularization $\epsilon$, the SK algorithm becomes extremely slow. To counter this, we propose a new algorithm that accelerates SK by \emph{overrelaxation}.
}

%----------------------------------------------------------------------------------------
%	Notations and definitions
%----------------------------------------------------------------------------------------

\headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0,below=intro}{
	We aim at numerically solving the entropic regularization of optimal transport:
	\begin{equation} \label{eq:problem}
	\gamma^* = \argmin_{\gamma \in \Ccal_1 \cap \Ccal_2}
	\scal{c}{\gamma} + \epsilon \KL(\gamma,\One).
	\end{equation}
	In this equation, $\Ccal_1$ and $\Ccal_2$ are the marginal constraint sets corresponding to two discrete measures $\mu^1$ and $\mu^2$,
	and where $\KL$ is the Kullback--Leibler divergence.
	\begin{gather*}
	\Ccal_1 = \left\{ \gamma \mid A_1 \gamma = \mu^1 \right\},
	\quad\quad\quad
	\Ccal_2 = \left\{ \gamma \mid A_2 \gamma = \mu^2 \right\}.\\
	\KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}.
	\end{gather*}
	\begin{minipage}{0.6\textwidth}
		As problem (\ref{eq:problem}) is strictly convex, it has exactly one solution. It can be written equivalently as the \emph{Bregman projection} of a point onto a convex set:
		\begin{gather*}
		P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} \KL(\gamma,\xi),\\
		\text{(\ref{eq:problem})} \quad \Longleftrightarrow \quad \gamma^* = P_{\Ccal_1 \cap \Ccal_2}(e^{-c/\epsilon}).
		\end{gather*}
		Whereas projecting onto $\Ccal_1 \cap \Ccal_2$ is challenging, projecting onto $\Ccal_1$ or $\Ccal_2$ individually is much easier. Indeed, it corresponds to scaling the rows or columns of the matrix so that the corresponding marginal sum is correct.
		\begin{align*}\label{scaling}
		P_{\Ccal_1}(\gamma) &= \diag(a) \gamma &\text{with}\quad
		a &=  {\mu^1}\oslash{A_1 \gamma} \\
		P_{\Ccal_2}(\gamma) &= \gamma \diag(b) &\text{with}\quad
		b &= {\mu^2}\oslash{A_2 \gamma}\nonumber
		\end{align*}
	\end{minipage}
	\begin{minipage}{0.39\textwidth}
		\begin{tcolorbox}[/tcb/size=small, /tcb/top=2.0mm]
		\centering
		\input{schema_a}
		\captionof{figure}{The Sinkhorn--Knopp algorithm iterates Bregman projections on the two constraint sets.}
		\end{tcolorbox}
	\end{minipage}

	As shown in [Bregman, 1964], the solution can be computed by alternating Bregman projections. The main idea of the Sinkhorn--Knopp algorithm can therefore be written as:
\begin{equation}
\lim P_{\Ccal_2}\circ P_{\Ccal_1} \circ \ldots \circ P_{\Ccal_2} \circ P_{\Ccal_1} (\gamma^0) = \gamma^*
\end{equation}
}

%----------------------------------------------------------------------------------------
%	
%----------------------------------------------------------------------------------------

\headerbox{Overrelaxation}{name=overrelaxation,column=1}{
Although the SK algorithm is much more efficient than linear solvers in a high-dimensional setting, it becomes extremely slow when the regularization $\epsilon$ is small.
To counteract this limitation, we use the principle of \emph{overrelaxation}.
\begin{minipage}{0.5\textwidth}
	\begin{tcolorbox}[size=small, top=2mm]
		\centering
		\begin{tabular}{c c}
		\input{schema_b} &
		\input{schema_c} \\
		SK &
		Overrelaxed SK
		\end{tabular}
	\captionof{figure}{For small regularization $\epsilon$, the SK algorithm converges slowly, whereas the overrelaxed SK algorithm converges with a better rate}
	\end{tcolorbox}
\end{minipage}
\begin{minipage}{0.49\textwidth}
	We define the overrelaxed projections of parameter $\theta \in \IR$:
	\begin{align*}\label{or_scaling}
	P_{\Ccal_1}^\theta(\gamma) &= \diag(a)^\theta \gamma\\
	P_{\Ccal_2}^\theta(\gamma) &= \gamma \diag(b)^\theta \nonumber
	\end{align*}
	(element-wise exponentiation)
\end{minipage}
}

%----------------------------------------------------------------------------------------
% Main result
%----------------------------------------------------------------------------------------

\headerbox{Lyapunov Function}{name=lyapunov,below=Sinkhorn,column=0}{

}

%----------------------------------------------------------------------------------------
%	Bootstrap
%----------------------------------------------------------------------------------------

\headerbox{Experiments}{name=expe,below=overrelaxation,column=1}{

}

%----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

\headerbox{References}{name=references,below=lyapunov,span =2}{
\noindent [1] M. Cuturi. Sinkhorn distances: Lightspeed computation of optimal transport. {\it Advances in Neural Information Processing Systems 26}, 2013.\\
...
}

\end{poster}

\end{document}